---
title: "A Sneak Peek to Chao Theory"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: flatly
---

<style>
.storyboard-nav .sbframelist, .storyboard-nav .sbnext, .storyboard-nav .sbprev {height: 77px !important;}
.shiny-input-container:not(.shiny-input-container-inline) {width: 90% !important;}
.shiny-split-layout>div{overflow: hidden;}
body {font-family: Georgia,"Times New Roman",Times,serif; color: #777}
h1 {font-size: xxlarge; text-align: center;}
p, li {font-size: larger; text-align: left;}
.sbframe-commentary {width: 66% !important;}

.dropdown-menu {position:fixed; top:inherit; left:inherit; max-width:360px}
.form-group {margin-bottom: 3px;}
.input-box {position: relative; }
.form-control {display: block; border: 1px solid #d7d6d6; background: #fff; padding: 10px 10px 10px 30px;}
.control-label{position: relative; display: block; left: 5px; top: 20px; z-index: 9; font-style: italic; height: 1px}
label[for=ptype], div.form-group > br{display:none}
</style>

```{r, echo=FALSE}
library(shiny)
library(shinyWidgets)
source("global.R")
knitr::opts_chunk$set(echo = F, message = F, warning = F)
#  add to ### to control size {data-commentary-width=550}
```



### Lorenz Equation Demo {data-commentary-width=800}

```{r lorenz_opts}
# autonumericInput(inputId="id3", label="", value = 1.0000, currencySymbol = "ρ=", decimalPlaces = 3)
splitLayout(cellWidths = c("70%","30%"), 
            NULL ,
            numericInput("tmax", "t=", min=1,  max=100, value=50,  step=1))
```


<div style="border-radius: 6px; border: 2px solid #fbb4ae; padding: 0px 9px; text-decoration: underline 3px #fbb4ae; margin: 9px 0px 9px">

Lorenz System 1

```{r}
splitLayout(cellWidths = c("33%", "33%", "33%"), 
            numericInput('rho1',   "ρ=", min=3, max=50, value=15, step=1),
            numericInput('sigma1', "σ=", min=3, max=50, value=10, step=1),
            numericInput('beta1',  "β=", min=3, max=50, value=2.67, step=1))
splitLayout(cellWidths = c("33%", "33%", "33%"), 
            numericInput("x01", "x₀=", min=0, max=10, value=1, step=0.001),
            numericInput("y01", "y₀=", min=0, max=10, value=1, step=0.001),
            numericInput("z01", "z₀=", min=0, max=10, value=1, step=0.001))
```

</div><div style="border-radius: 6px; border: 2px solid #b3cde3; padding: 0px 9px; text-decoration: underline 3px #b3cde3; margin-bottom: 9px;">

Lorenz System 2

```{r}
splitLayout(cellWidths = c("33%", "33%", "33%"), 
            numericInput('rho2',   "ρ=", min=3, max=50, value=28, step=1),
            numericInput('sigma2', "σ=", min=3, max=50, value=10, step=1),
            numericInput('beta2',  "β=", min=3, max=50, value=2.67, step=1))
splitLayout(cellWidths = c("33%", "33%", "33%"), 
            numericInput("x02", "x₀=", min=0, max=10, value=1.000, step=0.001),
            numericInput("y02", "y₀=", min=0, max=10, value=1.000, step=0.001),
            numericInput("z02", "z₀=", min=0, max=10, value=1.000, step=0.001))
```

</div>

```{r}
uiOutput("info")
```

***

```{r lorenz_layout}
smooth_disp = c("Fast", "Balanced", "Accurate", "Smooth")
smooth_value = c(50, 100, 200, 250)

splitLayout(
  cellWidths = c("25%","75%"),
  dropdownButton(
    materialSwitch("noAXES", "Hide Axes", status = "default", value = FALSE, right = TRUE),
    materialSwitch("show1",  "Show System #1", status = "default", value = T, right=TRUE),
    materialSwitch("show2",  "Show System #2", status = "default", value = T, right=TRUE),
    tags$hr(),
    tags$small("Render Quality:"),
    sliderTextInput(inputId = "smooth", label = " ", choices = smooth_disp,
                    selected = smooth_disp[2]),
    circle = F, label=" Options", status = "default",
    icon = icon("gear"), width = "99%"),
    
  radioGroupButtons(
    inputId = "ptype", label = "Plot Type",
    choices = c("<i class='fa fa-cubes'></i>  3D"="3d",
                "<i class='fa fa-braille'></i>  Projections"="prj", 
                "<i class='fa fa-line-chart'></i>  Time Sereis"="ts",
                "<i class='fa fa-route'></i>  Ttrajectories"="tr"),
    justified = TRUE)
)

plotlyOutput("the_plot", width="99%", height="85%")

```

```{r lorenz_server}
rptype <- reactive(input$ptype)

output$the_plot <- renderPlotly({
  
  the_freq <- 1/smooth_value[which(smooth_disp == input$smooth)]
  
  if(input$show1){
    df1 <-  lorenz(input$rho1, input$sigma1, input$beta1, input$x01, input$y01, input$z01,
                   input$tmax, the_freq, type="data",name="#fbb4ae")
  } else {
    df1 <- data.frame(x=1, y=1, z=1, t=0, name="#b3cde3")

  }

  if(input$show2){
    df2 <-  lorenz(input$rho2, input$sigma2, input$beta2, input$x02, input$y02, input$z02,
                   input$tmax, the_freq, type="data",name="#b3cde3")
  } else {
    df2 <- data.frame(x=1, y=1, z=1, t=0, name="#b3cde3")
  }
  
  output$info<- renderUI(tags$small(
    "Smoothness:", smooth_value[which(smooth_disp == input$smooth)], tags$br(),
    "Data entries: ", nrow(df1) + nrow(df2), tags$br()))

  
  if(rptype()=="3d"){
    fig <- rbind(df1, df2) %>%
      plot_ly() %>%
      add_trace(x=~x, y=~y, z=~z, color=~I(name), mode="lines", type="scatter3d") %>%
      layout(title = mytit)

    if(input$noAXES) return(fig %>% layout(scene=list(xaxis=noaxis, yaxis=noaxis, zaxis=noaxis)))
    else return(fig)
  }
  
  if(rptype()=="prj"){return(
    subplot(
      plot_ly() %>%
        add_trace(data=df1, x=~x, y=~y, color=~I(name), mode="lines", type="scatter") %>%
        add_trace(data=df2, x=~x, y=~y, color=~I(name), mode="lines", type="scatter"),
      plot_ly() %>%
        add_trace(data=df1, x=~y, y=~z, color=~I(name), mode="lines", type="scatter") %>%
        add_trace(data=df2, x=~y, y=~z, color=~I(name), mode="lines", type="scatter"),
      plot_ly() %>%
        add_trace(data=df1, x=~x, y=~z, color=~I(name), mode="lines", type="scatter") %>%
        add_trace(data=df2, x=~x, y=~z, color=~I(name), mode="lines", type="scatter"),
      shareX = F, shareY = F, titleX = T, titleY = T, margin = 0.05)  %>%
        layout(showlegend=FALSE, title = mytit)
  )}

  if(rptype()=="ts"){return(
    subplot(
      plot_ly() %>%
        add_trace(data=df1, x=~t, y=~x, color=~I(name), mode="lines", type="scatter") %>%
        add_trace(data=df2, x=~t, y=~x, color=~I(name), mode="lines", type="scatter"),
      plot_ly() %>%
        add_trace(data=df1, x=~t, y=~y, color=~I(name), mode="lines", type="scatter") %>%
        add_trace(data=df2, x=~t, y=~y, color=~I(name), mode="lines", type="scatter"),
      plot_ly() %>%
        add_trace(data=df1, x=~t, y=~z, color=~I(name), mode="lines", type="scatter") %>%
        add_trace(data=df2, x=~t, y=~z, color=~I(name), mode="lines", type="scatter"),
      shareX = F, shareY = F, titleX = T, titleY = T, margin = 0.05)  %>%
        layout(showlegend=FALSE, hovermode="x", title = mytit)
  )}
  
  if(rptype()=="tr"){return(
    plot_ly() %>% 
      # # 2D-projection
      # add_trace(data=df1, x=~x, y=~y, color=~I(name), mode="lines", type="scatter") %>%
      # add_trace(data=df2, x=~x, y=~y, color=~I(name), mode="lines", type="scatter") %>%
      # add_markers(data=df1 %>% filter(floor(10*t)==10*t), x=~x, y=~y, frame=~t, marker=mstyle("cyan")) %>%
      # add_markers(data=df2 %>% filter(floor(10*t)==10*t), x=~x, y=~y, frame=~t, marker=mstyle("yellow")) %>%
      # layout(showlegend=FALSE)
      # 3D full
      add_trace(data=df1, x=~x, y=~y, z=~z, color=~I(name), mode="lines", type="scatter3d") %>%
      add_trace(data=df2, x=~x, y=~y, z=~z, color=~I(name), mode="lines", type="scatter3d") %>%
      add_markers(data=df1 %>% filter(floor(10*t)==10*t), x=~x, y=~y, z=~z, frame=~t, marker=mstyle("cyan",6)) %>%
      add_markers(data=df2 %>% filter(floor(10*t)==10*t), x=~x, y=~y, z=~z, frame=~t, marker=mstyle("yellow",6)) %>%
      layout(showlegend=FALSE, title = mytit, 
             scene=list(xaxis=noaxis, yaxis=noaxis, zaxis=noaxis)) %>%
      animation_slider(currentvalue = list(prefix = "Time ", font = list(color="red"))) %>%
      animation_opts(100)
  )}
  
  if(rptype()=="all"){return( # slow, not in use
    plot_ly(rbind(df1, df2)) %>%
      add_markers(x=~x, y=~y, z=~z, color=~I(name), type="scatter3d", size=I(3)) %>%
      layout(updatemenus=dropmenu)
  )}
})
```


### Introduction

<h1> A Sneak Peak to Chaos Theory</h1>

### Lorenz System (Lorenz, 1963)

> Deterministic nonperiodic flow

The mathematical model for atmospheric convection, in the domain of fluid dynamic system.

An ordinal differential equation with parameters $\sigma, \rho$ and $\beta$:

$$
\begin{cases} 
\frac{dx}{dt} = \sigma (y-x) \\
\frac{dy}{dt} = x(\rho-z) -y \\
\frac{dz}{dt} = -\beta z + xy \\
\end{cases}
$$

or simply by logistic map (recurrent mapping of $x,y,z$)

$$
\begin{cases} 
x' \to \sigma (y-x) \\
y' \to x(\rho-z) -y \\
z' \to  -\beta z + x y \\
\end{cases}
$$


The original Lorenz System is defined by $\rho=28, \sigma=10$ and $\beta= 8/3$, and was derived from Lord Rayleigh’s thermal convection (a Fourier expansion of the stream function and temperature).

With the simple equations, the solutions are "beautifully chaotic":

- deterministic

- sensitive to initial values

- exists strange attractors

